\documentclass{article}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}   

\title{Documentation}
\begin{document}
\maketitle

\section{Ideas from Brady et al.}

Use species-area relationships?

Larger field sizes imply lower unit costs of production

\section{Initialization}
Among the 10\% bigger farms: 
\begin{itemize}
    \item 50\% of the plot are 100\% intensive
    \item 50\% of the plot are 90\% intensive and 10\% conservation.
\end{itemize}

Among the 50\% smaller farms: 
\begin{itemize}
    \item 50\% of the plot are 100\% intensive
    \item 50\% of the plot are 100\% organic.
\end{itemize}

Among the 40\% medium farms:
\begin{itemize}
    \item 30\% of the plot are 100\% intensive
    \item 30\% of the plot are 90\% intensive and 10\% conservation.
    \item 40\% of the plot are 100\% organic.
\end{itemize}

\section{Parameters \& Variables}

\subsection{Parameters}

\begin{table}[h!]
\centering
\begin{tabular}{|l|c|p{5cm}|}
\hline
\textbf{Parameter} & \textbf{Value Range} & \textbf{Description} \\
\hline
n\_rows & $> 0$ & Vertical size of the grid  \\
\hline
n\_cols & $> 0$ & Horizontal size of the grid  \\
\hline
(x,y) & $ (0,0)$ ; $(n_{rows}-1, n_{cols}-1) $ & Grid coordinates \\
\hline
PlotID & 0 ; $K - 1 = n_{rows} \times n_{cols} - 1$ & Plot Identifier \\
\hline
\hline
n\_farmers & 1 ; $N < K$ & Number of farmers \\
\hline
FarmerID & 0 ; $N - 1$ & Farmer Identifier \\
\hline
farm\_mu & $\mathbb{R}$ & Mean parameter of the underlying distribution (lognormal farm-size distribution) \\
\hline
farm\_sigma & $\geq 0$ & Standard deviation of the underlying distribution (lognormal farm-size distribution) \\
\hline
\hline
q & 0 ; 1 & Plot quality \\
\hline
q\_sigma & $\mathbb{R}$ & Scaling parameter before normalization. 

No change if $q > 0$, inverted map if $q < 0$, homogeneous quality if $q = 0$. \\
\hline
q\_length\_scale & $\geq 0$ & Length scale of the Gaussian kernel used to smooth the noise \\
\hline
\hline
alpha & $\mathbb{R}^+$ & Yield multiplicator\\
\hline
gamma & $\mathbb{R}^+$ & Yield exponent\\
\hline
c & $\mathbb{R}^+$ & Baseline cost  \\
\hline
kappa & $\mathbb{R}^+$ & Coefficient for quality-dependent cost \\
\hline
eta & $\mathbb{R}^+$ & Farmer-specific cost \\
\hline
output\_price & $\mathbb{R}^+$ & Price per unit of output. \\
\hline
\end{tabular}
\end{table}

\subsection{Variables}

\begin{table}[h!]
\centering
\begin{tabular}{|l|c|p{5cm}|}
\hline
\textbf{Variable} & \textbf{Value Range} & \textbf{Description} \\
\hline
LandUse & \{I, O, S\} & Land use: Intensive agriculture (I), Organic agriculture (O), or Conservation (S) \\
\hline
Yield & $\mathbb{R}^+$ & Yield of the plot \\
\hline
Cost & $\mathbb{R}^+$ & Production cost of the plot \\
\hline
Revenue & $\mathbb{R}^+$ & Production revenue of the plot, $\text{Price} \times \text{Yield}$ \\
\hline
profit\_before\_subsidy & $\mathbb{R}^+$ & Production profit of the plot, $\text{Revenue} - \text{Cost}$ \\
\hline
aggregate\_revenue & $\mathbb{R}^+$ & Total revenue of the farm, sum of the revenue of all plots owned by the farmer \\
\hline
aggregate\_profit & $\mathbb{R}^+$ & Total profit of the farm \\
\hline
\end{tabular}
\end{table}

\section{Modules}
\subsection{Landscape Generation}

The \emph{landscape} is a rectangular grid of $n_{rows} \times n_{cols}$, containing $K$ plots. \\

\noindent A \emph{plot} is defined by its ID \texttt{PlotID}, its coordinates \texttt{(x,y)}, and its quality \texttt{q}. \\
Von Neumann neighborhood (N, S, E, W) is used to determine the adjacent plots. \\
The Euclidean distance ($d = \sqrt{(x_2-x_1)^2 + (y_2-y_1)^2}$) is used to determine the distance between two plots. \\

\noindent The quality of each plot is generated from Gaussian noise with mean $0$, smoothed using a Gaussian kernel with length scale \texttt{q\_length\_scale}, then rescaled using \texttt{q\_sigma}. The quality values are then normalized on [0, 1]. \\

\noindent There are $N < K = n_{rows} \times n_{cols}$ \emph{farmers} in the landscape, each owning a number of plots drawn from a lognormal distribution with parameters \texttt{farm\_mu} and \texttt{farm\_sigma}. Each farmer owns at least one plot, and the total number of plots owned by all farmers equals $K$. 

\noindent Farmers are defined by their ID \texttt{FarmerID} and the list of plots \texttt{plotID} they own: the number of plots owned corresponds to the farm's size $m$. 

\subsection{Agriculture production}

Let us consider that a single crop is produced in the landscape, and that the price of the output is exogenous and constant over time. \\

\noindent The \emph{yield} of a plot depends on its quality $q$ and its land use (intensive agriculture, organic agriculture, or conservation), such that: 
\begin{equation}
    \text{Yield} = \alpha(\text{LandUse}) \cdot q^{\gamma(\text{LandUse})}
\end{equation}
where $\alpha$ and $\gamma$ are parameters that differ between land use types. \\

\noindent \emph{Costs} of production can be decomposed into a baseline cost (e.g. fixed costs of production), one that depends on plot quality, and a farmer-specific cost component drawn once at the beginning for each farmer from a normal distribution with mean $0$ and standard deviation $\sigma_{\eta}$.
\begin{equation}
    \text{Cost}_i = c(\text{LandUse}) + \kappa(\text{LandUse}) \cdot (1 - q) + \eta_i(\text{LandUse}), \quad \eta_i(\text{LandUse}) \sim \mathcal {N}(0,\sigma_\eta^2)
\end{equation}

\noindent The {revenue} is simply the product of yield and price, and the profit before subsidy is the difference between revenue and cost.


\end{document}